### 2024-11-05 manuscript.tex
Address comments from Graeme

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is the 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). Due to the resolution, parametrisations of convection and clouds are
avoided. Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Unstable conditions are the most problematic for the
models. In daily cloud cover, ICON and the reanalyses tend to be about 1 and 2
oktas clearer, respectively. Compared to radiosondes, potential temperature is
accurate in all, but ICON is too unstable over the low-latitude SO and too humid in
the boundary layer. MERRA-2 is too humid at all heights. SO cloud biases are a
substantial issue in the GSRM, but are an improvement over the
lower-resolution reanalyses. Explicitly resolved convection and cloud processes
were not enough to address the model cloud biases.

@@

Multiple cloud properties have an effect on the shortwave (SW) and longwave
(LW) radiation. To first order, the total cloud fraction, cloud
phase, and the liquid and ice water path are the most important cloud properties influencing SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. Second order effects on SW and LW radiation are associated with the cloud droplet size
distribution, ice crystal habit, cloud lifetime, and direct radiative
interaction with aerosols.  In the 6$^\mathrm{th}$ phase of the Coupled Model
Intercomparison Project [CMIP6; \cite{eyring2016}], the cloud feedback has
increased relative to CMIP5 \citep{zelinka2020}, this being one of the main
reasons for the higher climate sensitivity of CMIP6 models.

@@

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, due to, for example SST affecting
clouds and clouds affecting the surface radiation. \cite{zhang2023} has shown
that SST biases have improved in CMIP6 compared to CMIP5 [relative to the
European Centre for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5; ERA5],
with SST overall increasing in the later CMIP phase. However, over the SO this
resulted in an even higher positive bias, especially in the Atlantic Ocean (AO)
sector of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the
SO SST bias in an ensemble of 18 CMIP6 models originates not from the surface
heat and radiation fluxes (using reanalyses as a reference), but from a warm
bias in the Northern Atlantic Deep Water.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/lts_dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data available. Shown is also the
chosen dividing threshold of 12 K for the relatively stable and unstable
conditions.
}
\label{fig:lts}
\end{figure}

@@

In addition to the above, we partitioned our data into two mutually exclusive
subsets by stability. We determined this by calculating lower tropospheric
stability (LTS) as the difference between the potential temperature at 700 hPa
and the surface.  Based on a histogram of LTS in ERA5 and MERRA-2 calculated at
all voyage tracks and stations (Fig.  \ref{fig:lts}), we determined a dividing
threshold of 12 K for the relatively unstable (< 12 K) and the relatively
stable (>= 12 K) conditions.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
the relatively stable (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

Fig. \ref{fig:cyclone-stability}b shows a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}).  The relatively unstable conditions are prevalent
in the middle SO (50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. The relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 renalysis is also substantially more stable than
ICON across the whole region.

@@

In the general case (Fig. \ref{fig:cloud-occurrence}a), the observations show
cloud occurrence peaking at the surface, whereas models show a higher peak (at
about 500 m). The models underestimate the total cloud fraction by 10--20\% and
show a strong drop in cloud occurrence near the surface, but this is not
supported by the observations. ICON and ERA5 overestimate cloud occurrence at
the peak (between 0--1 km). Above 1 km, ICON and MERRA-2 understimate cloud
occurrence, but ERA5 is very accurate. The too high peak in models is partly
supported by the lifting condensation level (LCL) distribution, which peaks
higher in the models than in the observations (at the surface), although this
is not very pronounced.

When subsetted by low- and high-latitude zones (Fig.
\ref{fig:cloud-occurrence}b, c), we see that the low-latitude SO zone shows
a stronger peak of cloud occurrence near the surface than the high-latitude SO
zone, and this could be because higher latitudes have more unstable atmospheric
profiles. The low- and high-latitude SO zones show similar biases in models as
in the general case, but ERA5 does not overestimate the peak in the
low-latitude SO zone (near-surface cloud occurrence is still strongly
understimated).

@@

When subsetted by the relatively stable and unstable conditions (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in the relatively stable
situations cloud occurrence peaks strongly at the surface in observations,
compared to the relatively unstable situations, where the peak is more obtuse
and spread across the altitudes of 0--1 km.  In the relatively stable
situations, the models are doing a fairly good job, but overestimate cloud
occurrence at the peak below 1 km; above 1 km, they show similar biases as in
the general case.  In the relatively unstable situations, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 is
underestimating cloud occurrence below 1 km (especially near the surface), and
MERRA-2 is underestimating it even more strongly.

In all situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in the relatively unstable and
non-cyclonic conditions. It also shows a peak of cloud occurrence at higher
altitude than observations (500 m vs. near the surface), and correspondingly,
its LCL tends to be also higher. MERRA-2 can be generally characterised as
underestimating cloud occurrence at nearly all altitudes as well as the total
cloud fraction, but mostly above and below 500 m (the peak at 500 m is
well-represented). It struggles the most in the low-latitude SO zone and in the
relatively unstable situations. ERA5 can be generally characterised as
representing cloud occurrence correctly above about 1--1.5 km, overestimating
below, but underestimating near-surface cloud occurrence (0--500 m). The total
cloud fraction is strongly understimated in all situations. It has a tendency
towards understimation in the low-latitude SO zone and the relatively unstable
situations; conversely, overestimating in the high-latitude SO zone and the
relatively stable conditions.

@@

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in the previous section. Observations have the greatest representation of
high cloud cover (5--8 oktas), peaking at 7 oktas. This is not represented by
ICON or the reanalyses.  While ICON is the closest, it tends to be 1 okta
clearer the observations, peaking at 6 oktas, and highly understimating days
with 8 oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly understimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 shows slightly higher cloud
cover and thus is more consistent with observations.

When analysed by subsets, observations in the cyclonic subset show the highest
cloud cover, with 8 oktas occurring on one half of such days (Fig.
\ref{fig:cloud-cover}d).  This is not represented by ICON or the reanalyses at
all. Interestingly, clear sky days (0 oktas) also have a local maximum peaking
at about 15\% in this subset.  When we contrast the low- and high-latitude
zones, we see that the high-latitude zone tends to have greater cloud cover,
peaking at 8 oktas (Fig.  \ref{fig:cloud-cover}c). The high-latitude zone also
has almost no clear sky or small cloud cover cases (0--4 oktas). ICON and the
reanalyses represent at least this characteristic of the distribution well for
0--3 oktas, but otherwise show bises similar to the general case. One of the
greatest biases is present in ERA5 in the relatively unstable subset, in which
ERA5 peaks at 3 oktas, whereas the observations peak at 7 oktas and show
negligible cloud cover below 5 oktas.

@@

Overall, the mean $\theta_v$ is relatively well represented in ICON and
MERRA-2, being only slightly colder in the mid-to-high troposophere (less
stable) in ICON than in observations (Fig. \ref{fig:potential-temperature}a).
Large differences exist, however, in the 40--55°S zone, where ICON is colder in
$\theta_v$ by up to about 5 K and more so at higher altitudes (Fig.
\ref{fig:potential-temperature}b). In other subsets, the bias is relatively
small. MERRA-2 is very close to the observations, possibly due to a high
accuracy of assimilation of this quantity. Notably, the variability of virtual
potential temperature (as represented by the percentiles) is much smaller in
ICON than in the observations. This indicates that the model's internal
variability in the lower-tropospheric thermodynamic conditions in the SO is
smaller than in reality.

@@

Let us consider the main limitations of the presented results. The spatial
coverage of our dataset does not include most parts of the Indian Ocean and
Pacific Ocean sectors of the SO. Even though climatological features of the SO
are typically relatively uniform zonally, variantions exist, such as those
related to the Antarctic Peninsula and the southern tip of South America. The
voyages were mostly undertaken in the Austral summer months and only rarely in
the winter months, due to the poor accesibility of this region during winter.
Therefore, our results are mostly representative of summer and to a lesser
extent, spring and autumn conditions.

@@

We attempted to remove lidar profiles with precipitation, which could not be
properly simulated with the lidar simulator (Section \ref{sec:ann}). However,
the approach was limited by the relatively low sensitivity of the ANN (65\%)
and the fact that we had to choose a fixed threshold for surface precipitation
flux in the model and reanalyses, which might not exactly correspond to
detection by the ANN applied to obseravations. Also, we did not make an attempt
to remove profiles with precipitation not reaching the surface. The above
reasons can result in an artificial bias in the comparison, even though we
expect this to be much smaller than the identified model biases.

@@

We show that relative to ERA5, the distribution and strength of cyclonic
activity over the SO is well represented in ICON, but it is substantially less
stable in terms of LTS. The latter is also manifested in the radiosonde profile
comparison, which shows that the virtual potential temperature profiles in ICON
are more stable than in the observations over low-latitude SO.

The 31 voyages and a station show fairly similar biases in cloud occurrence by
height in the lidar comparison, which indicates that common underlying causes
for the biases exist regardless of longitude and season. ICON underestimates
the total cloud fraction by about 10\%, with an overestimation of clouds below
2 km and an underestimation of clouds above 2 km. The reanalyses also
underestimate the total cloud fraction by about 20\%.  ERA5 overestimates cloud
below 1 km but underestimates near-surface cloud or fog. ICON strongly
overestimates the peak of cloud occurrence at about 500 m, which might be
explained by the radiosonde comparison, which shows that it is too moist at
around this height.  Similar to our results, \cite{cesana2022} showed that
CMIP6 models also tend to underestimate cloud occurrence above 2 km over the
SO, although their analysis in this case was limited to liquid clouds.

Compared to the lidar observations, the daily cloud cover tends to be about 1
okta lower in ICON and 2 oktas lower in the reanalyses.  Unstable conditions
are associated with some of the greatest biases, especially in ERA5.  The
models also underestimate the cloud cover very strongly in cyclonic conditions,
which are very cloudy in the observations (8 oktas), but much less so in the
models.

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is the 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). Due to the high resolution, parametrisations of convection and clouds
are avoided. Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Unstable conditions are the most problematic for the
models. In daily cloud cover, ICON and the reanalyses tend to be about 1 and 2
oktas clearer, respectively. Compared to radiosondes, potential temperature is
accurate in all, but ICON is too unstable over the low-latitude SO and too
humid in the boundary layer. MERRA-2 is too humid at all heights. SO cloud
biases are a substantial issue in the GSRM, but are an improvement over the
lower-resolution reanalyses. Explicitly resolved convection and cloud processes
were not enough to address the model cloud biases.

@@

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW)
radiation. To first order, the total cloud fraction, cloud phase, and the
liquid and ice water path are the most important cloud properties influencing
SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. Second order effects on SW and LW radiation are associated
with the cloud droplet size distribution, ice crystal habit, cloud lifetime,
and direct radiative interaction with aerosols.  In the 6$^\mathrm{th}$ phase
of the Coupled Model Intercomparison Project [CMIP6; \cite{eyring2016}], the
cloud feedback has increased relative to CMIP5 \citep{zelinka2020}, this being
one of the main reasons for the higher climate sensitivity of CMIP6 models.

@@

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, for example SST affecting clouds and
clouds affecting the surface radiation. \cite{zhang2023} has shown that SST
biases have improved in CMIP6 compared to CMIP5 [relative to the European
Centre for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5; ERA5], with SST
overall increasing in the later CMIP phase. However, over the SO this resulted
in an even higher positive bias, especially in the Atlantic Ocean (AO) sector
of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the SO SST
bias in an ensemble of 18 CMIP6 models originates not from the surface heat and
radiation fluxes (using reanalyses as a reference), but from a warm bias in the
Northern Atlantic Deep Water.

@@

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/lts_dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data available. Shown is also the
chosen dividing threshold of 12 K for relatively stable and unstable
conditions.
}
\label{fig:lts}
\end{figure}

@@

In addition to the above, we partitioned our data into two mutually exclusive
subsets by stability. We determined this by calculating lower tropospheric
stability (LTS) as the difference between the potential temperature at 700 hPa
and the surface.  Based on a histogram of LTS in ERA5 and MERRA-2 calculated at
all voyage tracks and stations (Fig.  \ref{fig:lts}), we determined a dividing
threshold of 12 K for relatively unstable (< 12 K) and relatively stable (>= 12
K) conditions.

@@

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
relatively stable (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

@@

Fig. \ref{fig:cyclone-stability}b shows a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}). Relatively unstable conditions are prevalent
in the middle SO (50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. Relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 renalysis is also substantially more stable than
ICON across the whole region.

@@

In the general case (Fig. \ref{fig:cloud-occurrence}a), the observations show
cloud occurrence peaking at the surface, whereas models show a higher peak (at
about 500 m). The models underestimate the total cloud fraction by 10--20\% and
show a strong drop in cloud occurrence near the surface, but this is not
supported by the observations. ICON and ERA5 overestimate cloud occurrence at
the peak (between 0--1 km). Above 1 km, ICON and MERRA-2 underestimate cloud
occurrence, but ERA5 is very accurate. The exaggerated peak in models is partly
supported by the lifting condensation level (LCL) distribution, which peaks
higher in the models than in the observations (at the surface), although this
is not very pronounced.

When subsetted by low- and high-latitude zones (Fig.
\ref{fig:cloud-occurrence}b, c), we see that the low-latitude SO zone shows
a stronger peak of cloud occurrence near the surface than the high-latitude SO
zone, and this could be because higher latitudes have more unstable atmospheric
profiles. The low- and high-latitude SO zones show similar biases in models as
in the general case, but ERA5 does not overestimate the peak in the
low-latitude SO zone (near-surface cloud occurrence is still strongly
underestimated).

@@

When subsetted by relatively stable and unstable conditions (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in relatively stable
situations cloud occurrence peaks strongly at the surface in observations,
compared to relatively unstable situations, where the peak is more obtuse
and spread across the altitudes of 0--1 km.  In relatively stable
situations, the models are doing a fairly good job, but overestimate cloud
occurrence at the peak below 1 km; above 1 km, they show similar biases as in
the general case.  In relatively unstable situations, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 is
underestimating cloud occurrence below 1 km (especially near the surface), and
MERRA-2 is underestimating it even more strongly.

In all situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in relatively unstable and
non-cyclonic conditions. It also shows a peak of cloud occurrence at higher
altitude than observations (500 m vs. near the surface), and correspondingly,
its LCL tends to be also higher. MERRA-2 can be generally characterised as
underestimating cloud occurrence at nearly all altitudes as well as the total
cloud fraction, but mostly above and below 500 m (the peak at 500 m is
well-represented). It struggles the most in the low-latitude SO zone and in the
relatively unstable situations. ERA5 can be generally characterised as
representing cloud occurrence correctly above about 1--1.5 km, overestimating
below, but underestimating near-surface cloud occurrence (0--500 m). The total
cloud fraction is strongly underestimated in all situations. It has a tendency
towards underestimation in the low-latitude SO zone and relatively unstable
situations; conversely, overestimating in the high-latitude SO zone and the
relatively stable conditions.

@@

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in the previous section. Observations have the greatest representation of
high cloud cover (5--8 oktas), peaking at 7 oktas. This is not represented by
ICON or the reanalyses.  While ICON is the closest, it tends to be 1 okta
clearer the observations, peaking at 6 oktas, and highly underestimating days
with 8 oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly underestimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 shows slightly higher cloud
cover and thus is more consistent with observations.

When analysed by subsets, observations in the cyclonic subset show the highest
cloud cover, with 8 oktas occurring on one half of such days (Fig.
\ref{fig:cloud-cover}d).  This is not represented by ICON or the reanalyses at
all. Interestingly, clear sky days (0 oktas) also have a local maximum peaking
at about 15\% in this subset.  When we contrast the low- and high-latitude
zones, we see that the high-latitude zone tends to have greater cloud cover,
peaking at 8 oktas (Fig.  \ref{fig:cloud-cover}c). The high-latitude zone also
has almost no clear sky or small cloud cover cases (0--4 oktas). ICON and the
reanalyses represent at least this characteristic of the distribution well for
0--3 oktas, but otherwise show biases similar to the general case. One of the
greatest biases is present in ERA5 in the relatively unstable subset, in which
ERA5 peaks at 3 oktas, whereas the observations peak at 7 oktas and show
negligible cloud cover below 5 oktas.

@@

Overall, the mean $\theta_v$ is relatively well represented in ICON and
MERRA-2, being only slightly colder in the mid-to-high troposphere (less
stable) in ICON than in observations (Fig. \ref{fig:potential-temperature}a).
Large differences exist, however, in the 40--55°S zone, where ICON is colder in
$\theta_v$ by up to about 5 K and more so at higher altitudes (Fig.
\ref{fig:potential-temperature}b). In other subsets, the bias is relatively
small. MERRA-2 is very close to the observations, possibly due to a high
accuracy of assimilation of this quantity. Notably, the variability of virtual
potential temperature (as represented by the percentiles) is much smaller in
ICON than in the observations. This indicates that the model's internal
variability in the lower-tropospheric thermodynamic conditions in the SO is
smaller than in reality.

@@

Let us consider the main limitations of the presented results. The spatial
coverage of our dataset does not include most parts of the Indian Ocean and
Pacific Ocean sectors of the SO. Even though climatological features of the SO
are typically relatively uniform zonally, variations exist, such as those
related to the Antarctic Peninsula and the southern tip of South America. The
voyages were mostly undertaken in the Austral summer months and only rarely in
the winter months, due to the poor accesibility of this region during winter.
Therefore, our results are mostly representative of summer and to a lesser
extent, spring and autumn conditions.

@@

We attempted to remove lidar profiles with precipitation, which could not be
properly simulated with the lidar simulator (Section \ref{sec:ann}). However,
the approach was limited by the relatively low sensitivity of the ANN (65\%)
and the fact that we had to choose a fixed threshold for surface precipitation
flux in the model and reanalyses, which might not exactly correspond to
detection by the ANN applied to observations. Also, we did not make an attempt
to remove profiles with precipitation not reaching the surface. The above
reasons can result in an artificial bias in the comparison, even though we
expect this to be much smaller than the identified model biases.

@@

We show that relative to ERA5, the distribution and strength of cyclonic
activity over the SO is well represented in ICON, but it is substantially less
stable in terms of LTS. The latter is also manifested in the radiosonde profile
comparison, showing that the virtual potential temperature profiles in ICON
are more stable than in the observations over low-latitude SO.

The 31 voyages and a station show fairly similar biases in cloud occurrence by
height in the lidar comparison, which indicates that common underlying causes
for the biases exist regardless of longitude and season. ICON underestimates
the total cloud fraction by about 10\%, with an overestimation of clouds below
2 km and an underestimation of clouds above 2 km. The reanalyses also
underestimate the total cloud fraction by about 20\%.  ERA5 overestimates cloud
below 1 km but underestimates near-surface cloud or fog. ICON strongly
overestimates the peak of cloud occurrence at about 500 m, which might be
explained by the radiosonde comparison, showing that it is too moist at
around this height.  Similar to our results, \cite{cesana2022} showed that
CMIP6 models also tend to underestimate cloud occurrence above 2 km over the
SO, although their analysis in this case was limited to liquid clouds.

Compared to lidar observations, the daily cloud cover tends to be about 1 okta
lower in ICON and 2 oktas lower in the reanalyses.  Unstable conditions are
associated with some of the greatest biases, especially in ERA5.  The models
also underestimate the cloud cover very strongly in cyclonic conditions, which
are very cloudy in the observations (8 oktas), but much less so in the models.
